What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-09 lda-paper.tex
some edits but more data to look at

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
carefully examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
``diffs'', ``annotate'' and ``history'', these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word ``bitkeeper'' appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. ``RENAME'' also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with onlyy a little familiarity with the code base that
naming the topic was straightforward.

@@

We tracked Firebird from August 2000 to January 2006.
Our first attempt with Firebird resulted in a set of trends that were
dominated by one large trend containing tokens that were mostly stop
words such as ``the to a for of is it'' and references to one author
\texttt{carlosga05}.

@@

The third largest trend contained references to the changelog, the
build system, architecture and platform support. It specifically
mentioned tokens such as ``Solaris port x86''.

@@

Topics that were not trends but appeared to be  interesting were mostly external bug
fixes submitted to the project.
 In these cases, the developers would express thanks in their changelog comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
 tokens and topics such as: ``compiler workarounds'', ``nightly updates'', packets and MSVC++,

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs were common during some months. For instance in the
last month of MaxDB 7.500 development every topic contained 1 unique
Bug tracker URL, this pattern did not occur in the previou month. We
investigated the actual revisions and we did find there was difference
in behaviour between and that the last month of development commits
were mostly referencing the bug tracker. If the topics of one month
were 20 or more unique bug tickets being addressed, the global LDA
analysis would probably miss this, yet these bug tickets were the
focus of development for that month and not necessarily globally
relevant.

>>>>

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
carefully examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with onlyy a little familiarity with the code base that
naming the topic was straightforward.

@@

We tracked Firebird from August 2000 to January 2006.
Our first attempt with Firebird resulted in a set of trends that were
dominated by one large trend containing tokens that were mostly stop
words such as references to one author
\texttt{carlosga05} and the words: \emph{the, to, a, for, of, is, it}.

@@

The third largest trend contained references to the changelog, the
build system, architecture and platform support. It specifically
mentioned tokens such as \emph{Solaris}, \emph{port} and \emph{x86}.

@@

Topics that were not trends but appeared to be  interesting were mostly external bug
fixes submitted to the project.
 In these cases, the developers would express thanks in their changelog comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
 tokens and topics such as: \emph{compiler workarounds}, \emph{nightly updates}, \emph{packets} and \emph{MSVC++}.

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs dominated unique topics during some months. For
instance in the last month of MaxDB 7.500 development every topic
contained 1 unique Bug tracker URL, this pattern did not occur in the
previou month. We investigated the actual revisions and we did find
there was difference in behaviour between and that the last month of
development commits were mostly referencing the bug tracker. If the
topics of one month were 20 or more unique bug tickets being
addressed, the global LDA analysis would probably miss this, yet these
bug tickets were the focus of development for that month and not
necessarily globally relevant.

We noticed that Perforce and Changelist were in just about every
topic, so we added them to the stop words list. Longer trends were
shortened and in some cases the total number of topics found per month
was less. Evaluating different similarity thresholds showed that by
removing common words one reduces the general similarity of topics,
that said the larger topics still existed. Thus if you add more
relevant stop words one should tune your topic similarity parameters
to handle this change.
